---
title: "REM analysis"

---

```{r}

source('setup.R')

```


Load the data: 

```{r}

library('openxlsx')
library('dplyr')

path.REM <- '../../Data/REM/'
path.data <- '../../Data/data/'

data.visit1 <- read.xlsx(paste0(path.data,'Copy_of_GN_data_Visit1.xlsx'),sheet = 1)
data.visit2 <- read.xlsx(paste0(path.data,'Copy_of_GN_data_Visit2.xlsx'),sheet = 1)

data.rem1 <- na.omit(read.csv(paste0(path.REM,'REIGandTarget65_ver3.csv')))
data.rem2 <- na.omit(read.xlsx(paste0(path.REM,'REIG_Combined_Ver5.xlsx'),sheet = 1))


colnames(ssq_diff)[1] <- c('RC_ID')
colnames(df_ioi)[1] <- c('RC_ID')
colnames(df_audiogram_left)[10] <- c('RC_ID')
colnames(df_audiogram_right)[10] <- c('RC_ID')


```
Merge datasets: 



Compare REIG targets to REM measurements in plot

```{r}
library('dplyr')
library('ggplot2')
library('reshape2')
library('gridExtra')

REIG.measured.left <- data.rem1 %>% select(RC_ID,REIG250L, REIG500L, REIG1000L, REIG2000L, REIG4000L, REIG8000L)
REIG.measured.right <- data.rem1 %>% select(RC_ID,REIG250R, REIG500R, REIG1000R, REIG2000R, REIG4000R, REIG8000R)

REIG.target.left <- data.rem1 %>% select(RC_ID,Target250L, Target500L, Target1000L, Target2000L, Target4000L, Target8000L)
REIG.target.right <- data.rem1 %>% select(RC_ID,Target250R, Target500R, Target1000R, Target2000R, Target4000R, Target8000R)

diff.left <- REIG.measured.left
diff.right <- REIG.measured.right

diff.left[,2:7] <- REIG.measured.left[,2:7] - REIG.target.left[,2:7]
diff.right[,2:7] <- REIG.measured.right[,2:7] - REIG.target.right[,2:7]


diff.left.plot <- diff.left %>% melt()
diff.right.plot <- diff.right %>% melt()

png('Deviation_target.png',width = 1000, height = 600)

p1 <- ggplot(diff.left.plot, aes(variable,value)) + 
  geom_boxplot(fill = 'lightblue', color = 'blue', outlier.alpha = 0.1, width = .4) + 
  xlab('') + ylab('Measured - target gain (dB)') + labs(title = 'Left ear') + 
  theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16)) +
  scale_x_discrete(labels = c('250','500','1000','2000','4000','8000'))
  
p2 <- ggplot(diff.right.plot, aes(variable,value)) + 
  geom_boxplot(fill = 'lightblue', color = 'blue', outlier.alpha = 0.1, width = .4) + 
  xlab('') + ylab('') +labs(title = 'Right ear') +
   theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16)) +
  scale_x_discrete(labels = c('250','500','1000','2000','4000','8000'))
  

grid.arrange(p1,p2,ncol = 2)

dev.off()


```

Now cluster subjects into differences for both ears

```{r}
library('stats')
library('plyr')

diff <- na.omit(merge(diff.left, diff.right, by=c('RC_ID')))


# Determine number of clusters
wss <- (nrow(diff)-1)*sum(apply(diff[,2:13],2,var))
for (i in 2:15) wss[i] <- sum(kmeans(diff[,2:13],
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")

#lets go for 5 cluster centers

```

Do actual clustering

```{r}

knr <- 4

fit.kmeans <- kmeans(diff[,2:13],knr,nstart = 20)

means.left <- as.data.frame(fit.kmeans$centers) %>% select(REIG250L, REIG500L, REIG1000L, REIG2000L, REIG4000L, REIG8000L)
means.right <- as.data.frame(fit.kmeans$centers) %>% select(REIG250R, REIG500R, REIG1000R, REIG2000R, REIG4000R, REIG8000R)


means.left.plot <- means.left %>% data.frame(seq(1,knr,1)) 
means.right.plot <- means.right %>% data.frame(seq(1,knr,1)) 

#colnames(means.left.plot) <-c(colnames(means.left),'cluster')
#colnames(means.right.plot) <-c(colnames(means.right),'cluster')

colnames(means.left.plot) <- c('250','500', '1000', '2000', '4000', '8000','cluster')
colnames(means.right.plot) <- c('250','500', '1000', '2000', '4000', '8000','cluster')

p1 <- means.left.plot %>% melt(id=c('cluster')) %>% ggplot(aes(x = variable, y = value, color = factor(cluster),group=factor(cluster))) +
  geom_point(size=3) +
  geom_line(size=1.5) +
  xlab('') + ylab('Difference (dB)') + 
  scale_y_continuous(limits= c(-60,10)) + 
  theme(legend.position = 'none') + 
  xlab('Frequency') + ylab('Difference (dB)') + 
   theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16))
  

p2 <- means.right.plot %>% melt(id=c('cluster')) %>% ggplot(aes(x = variable, y = value, color=factor(cluster),group=factor(cluster))) +
  geom_point(size=4) + 
  geom_line(size=2) +
  xlab('') + ylab('') +
  scale_y_continuous(limits= c(-60,10)) + 
  scale_color_discrete(name = 'Cluster',
                      labels = c(paste0('1 ','(',as.character(fit.kmeans$size[1]),')'),
                                 paste0('2 ','(',as.character(fit.kmeans$size[2]),')'),
                                 paste0('3 ','(',as.character(fit.kmeans$size[3]),')'),
                                 paste0('4 ','(',as.character(fit.kmeans$size[4]),')'))) + 
  xlab('Frequency') + ylab('Difference (dB)') + 
   theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16))

  
png('Cluster_REM_bil.png',width = 1200, height = 700)

grid.arrange(p1,p2,ncol = 2)

dev.off()

```
Add cluster class to data

```{r}

data.rem1.cluster <- data.rem1 %>% data.frame(fit.kmeans$cluster) %>% distinct()

```


Connect to IOI

```{r}
#merge with IOI
data.merged.ioi <- df_ioi %>% merge(data.rem1.cluster, by = c('RC_ID')) %>% select('RC_ID','ioi_ha_1','ioi_ha_2','ioi_ha_3','ioi_ha_4','ioi_ha_5','ioi_ha_6',
                                                                           'ioi_ha_7','fit.kmeans.cluster')

data.ioi.plot <-  na.omit(data.merged.ioi) %>%  melt(id = c('RC_ID','fit.kmeans.cluster'))
data.ioi.plot$value <- data.ioi.plot$value %>% as.numeric()

data.ioi.agg <- aggregate(data.ioi.plot,by=list(data.ioi.plot$variable,data.ioi.plot$fit.kmeans.cluster),FUN=mean)

png('REM_IOI_bil.png',width = 900, height = 600)

ggplot(data.ioi.agg, aes(x=Group.1, y=value, fill = factor(Group.2))) + 
  geom_bar(stat = 'identity',position = 'dodge')+ xlab('Item #') + ylab('Value')+
  scale_fill_discrete(name = 'Cluster'
                       ) + 
  theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16))

dev.off()

```

Connect to SSQ difference

```{r}
#use %in% to make new data frame 

ssq.base <- df_ssq_base %>% filter(record_id %in% df_ssq_follow$record_id) 
ssq.diff <- df_ssq_follow[,-c(2)] #remove visit since we dont need that anymore
ssq.diff[,2:18] <- ssq.diff[,2:18] - ssq.base[,3:19] 
colnames(ssq.diff)[1] <- c('RC_ID')

#now merge with rem 

data.merged.ssq <- ssq.diff %>% merge(data.rem1.cluster, by = c('RC_ID')) %>% select(c(colnames(data.merged.ssq)[1:18],'fit.kmeans.cluster')) %>% distinct()
data.ssq.plot <-  na.omit(data.merged.ssq) %>%  melt(id = c('RC_ID','fit.kmeans.cluster'))

png('REM_SSQ_difference_bil.png',width = 1600, height = 900)

ggplot(data.ssq.plot, aes(x=variable, y=value, fill = factor(fit.kmeans.cluster))) + 
  geom_boxplot() + xlab('Item #') + ylab('Difference 1st to 2nd visit') + 
  scale_fill_discrete(name = 'Cluster'
                       ) + 
  theme(axis.title.y =  element_text(size = 16),
        axis.title.x = element_text(size = 16),
        legend.title = element_text(size = 16),
        text = element_text(size=12))

dev.off()


```
Now connect to audiogram

```{r}


data.aud.right <- read.csv(paste0(path.data,'Audiogram_right_classified.csv'))
data.aud.left <- read.csv(paste0(path.data,'Audiogram_left_classified.csv'))

colnames(data.aud.right)[which(colnames(data.aud.right) == 'ID')] <- c('RC_ID')
colnames(data.aud.left)[which(colnames(data.aud.left) == 'ID')] <- c('RC_ID')

#merge right side
data.rem.aud.ssq.right <- data.merged.ssq %>% filter(RC_ID %in% data.aud.right$RC_ID) %>%
  merge(data.aud.right, by = c('RC_ID'))

data.rem.aud.ssq.left <- data.merged.ssq %>% filter(RC_ID %in% data.aud.left$RC_ID) %>%
  merge(data.aud.left, by = c('RC_ID')) 


data.rem.aud.ssq.right <- data.rem.aud.ssq.right %>% data.frame('EAR'=c('right'))
data.rem.aud.ssq.left <- data.rem.aud.ssq.left %>% data.frame('EAR'=c('left'))


colnames(data.rem.aud.ssq.left)[which(colnames(data.rem.aud.ssq.left) == 'Age')] <- c('AGE')


data.rem.aud.ssq <- rbind(data.rem.aud.ssq.left,data.rem.aud.ssq.right)  %>% distinct()

data.rem.aud.ssq.plot <- data.rem.aud.ssq  %>% select('AGE','fit.kmeans.cluster','Class','Gender','RC_ID',colnames(data.rem.aud.ssq)[2:18]) %>% na.omit() %>% melt(id = c('RC_ID','fit.kmeans.cluster','Class','AGE','Gender'))

#data.rem.aud.ssq.plot <-  na.omit(data.rem.aud.ssq) %>%  melt(id = c('RC_ID','fit.kmeans.cluster','Class','AGE','Gender'))
data.rem.aud.ssq.plot$value <- data.rem.aud.ssq.plot$value  %>% as.numeric()

png('SSQ_REM_aud_bil.png',width = 2500, 1400)

ggplot(data.rem.aud.ssq.plot, aes(x=variable, y=value, fill = factor(fit.kmeans.cluster))) + 
  geom_boxplot() + xlab('Item #') + ylab('Difference 1st to 2nd visit') + 
  scale_fill_discrete(name = 'Cluster'
                       ) + 
  theme(axis.title.y =  element_text(size = 16),
        axis.title.x = element_text(size = 16),
        legend.title = element_text(size = 16),
        text = element_text(size=12))  + 
  facet_wrap(~Class, ncol = 5)
dev.off()

```

Now let's connect to the draw users (bilateral)


```{r}
tmp1 <- data.aud.cluster.ssq %>% select(colnames(data.aud.cluster.ssq)[9:25],'RC_ID','ClusterAud','AGE','Gender','Class','EAR')

data.draw.cluster <- t3.data.ssq %>% select('RC_ID','IsT1DrawerUser') %>% merge(tmp1, by = c('RC_ID')) %>% distinct()

data.draw.cluster <- transform(data.draw.cluster, SSQmean = rowMeans(data.draw.cluster[,3:19], na.rm = TRUE))

library('rpart')
library('partykit')

fit.tree <- rpart(as.factor(ClusterAud) ~ SSQmean + Class + Gender + AGE, data = data.draw.cluster, control = rpart.control(cp = 0.01))

png('Tree_REM.png', width = 1000, height = 800)

plot(as.party(fit.tree))

dev.off()

```





#####################################################################################################################################
Do clustering only for one side

```{r}

library('dplyr')

REIG.measured.left <- data.rem1 %>% select(RC_ID,REIG250L, REIG500L, REIG1000L, REIG2000L, REIG4000L, REIG8000L)
REIG.measured.right <- data.rem1 %>% select(RC_ID,REIG250R, REIG500R, REIG1000R, REIG2000R, REIG4000R, REIG8000R)

REIG.target.left <- data.rem1 %>% select(RC_ID,Target250L, Target500L, Target1000L, Target2000L, Target4000L, Target8000L)
REIG.target.right <- data.rem1 %>% select(RC_ID,Target250R, Target500R, Target1000R, Target2000R, Target4000R, Target8000R)

diff.left <- REIG.measured.left
diff.right <- REIG.measured.right

diff.left[,2:7] <- REIG.measured.left[,2:7] - REIG.target.left[,2:7]
diff.right[,2:7] <- REIG.measured.right[,2:7] - REIG.target.right[,2:7]

colnames(diff.left) <- c("RC_ID",   "250"  ,"500" , "1000" ,"2000", "4000" ,"8000")
colnames(diff.right) <- c("RC_ID",  "250"  ,"500" , "1000" ,"2000", "4000" ,"8000")

diff <- rbind(diff.right, diff.left) %>% select("250"  ,"500" , "1000" ,"2000", "4000" ,"8000")


```


```{r}
# Determine number of clusters
wss <- (nrow(diff)-1)*sum(apply(diff,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(diff,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")

#lets go for 3 cluster centers?


```

```{r}

library('reshape2')
library('ggplot2')
library('gridExtra')

knr <- 5

fit.kmeans <- kmeans(diff,knr,nstart = 20)

#means <- as.data.frame(fit.kmeans$centers) %>% select('250','500', '1000', '2000', '4000', '8000')

means.plot <- fit.kmeans$centers %>% data.frame(seq(1,knr,1)) 

colnames(means.plot) <- c('250','500', '1000', '2000', '4000', '8000','cluster')

png('Cluster_REM.png',width = 800, 
    height = 500)

p2 <- means.plot %>% melt(id=c('cluster')) %>% ggplot(aes(x = variable, y = value, color=factor(cluster),group=factor(cluster))) +
  geom_point(size=6) + 
  geom_line(size=2) +
  xlab('') + ylab('') +
  scale_y_continuous(limits= c(-60,10)) + 
  scale_color_discrete(name = 'Cluster',
                      labels = c(paste0('1 ','(',as.character(fit.kmeans$size[1]),')'),
                                 paste0('2 ','(',as.character(fit.kmeans$size[2]),')'),
                                 paste0('3 ','(',as.character(fit.kmeans$size[3]),')'),
                                 paste0('4 ','(',as.character(fit.kmeans$size[4]),')'),
                                 paste0('5 ','(',as.character(fit.kmeans$size[5]),')'))) + 
  xlab('Frequency') + ylab('Difference (dB)') + 
   theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16))



grid.arrange(p2,ncol = 1)

dev.off()

```

```{r}

data.rem1.cluster <- diff %>% data.frame(fit.kmeans$cluster) %>% data.frame(rbind(diff.left[1],diff.right[1])) 
colnames(df_ioi)[which(colnames(df_ioi) == 'record_id')] <- c('RC_ID')

```


```{r}

data.merged.ioi <- df_ioi %>% merge(data.rem1.cluster, by = c('RC_ID')) %>% select('RC_ID','ioi_ha_1','ioi_ha_2','ioi_ha_3','ioi_ha_4','ioi_ha_5','ioi_ha_6',
                                                                           'ioi_ha_7','fit.kmeans.cluster') %>% distinct()

data.ioi.plot <-  na.omit(data.merged.ioi) %>%  melt(id = c('RC_ID','fit.kmeans.cluster'))
data.ioi.plot$value <- data.ioi.plot$value %>% as.numeric()

data.ioi.agg <- aggregate(data.ioi.plot,by=list(data.ioi.plot$variable,data.ioi.plot$fit.kmeans.cluster),FUN=mean)

png('REM_IOI.png', width= 1000, height = 500)

ggplot(data.ioi.agg, aes(x=Group.1, y=value, fill = factor(Group.2))) + 
  geom_bar(stat = 'identity',position = 'dodge') + xlab('Item #') + ylab('Value')+
  scale_fill_discrete(name = 'Cluster'
                       ) + 
  theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16))

  
dev.off()


```




```{r}
#use %in% to make new data frame 

ssq.base <- df_ssq_base %>% filter(record_id %in% df_ssq_follow$record_id) 
ssq.diff <- df_ssq_follow[,-c(2)] #remove visit since we dont need that anymore
ssq.diff[,2:18] <- ssq.diff[,2:18] - ssq.base[,3:19] 
colnames(ssq.diff)[1] <- c('RC_ID')

#now merge with rem 
data.merged.ssq <- ssq.diff %>% merge(data.rem1.cluster, by = c('RC_ID')) %>% select(c(colnames(data.merged.ssq)[1:18],'fit.kmeans.cluster')) %>% distinct()

data.ssq.plot <-  na.omit(data.merged.ssq) %>%  melt(id = c('RC_ID','fit.kmeans.cluster'))

png('REM_SSQ_difference.png',width = 1300,height = 600)

ggplot(data.ssq.plot, aes(x=variable, y=value, fill = factor(fit.kmeans.cluster))) + 
  geom_boxplot() + xlab('Item #') + ylab('Difference 1st to 2nd visit') + 
  scale_fill_discrete(name = 'Cluster'
                       ) + 
  theme(axis.title.y =  element_text(size = 16),
        axis.title.x = element_text(size = 16),
        legend.title = element_text(size = 16),
        text = element_text(size=12))

dev.off()

```

Connect to audiogram 

```{r}

data.aud.right <- read.csv(paste0(path.data,'Audiogram_right_classified.csv'))
data.aud.left <- read.csv(paste0(path.data,'Audiogram_left_classified.csv'))

colnames(data.aud.right)[which(colnames(data.aud.right) == 'ID')] <- c('RC_ID')
colnames(data.aud.left)[which(colnames(data.aud.left) == 'ID')] <- c('RC_ID')

#merge right side
data.rem.aud.ssq.right <- data.merged.ssq %>% filter(RC_ID %in% data.aud.right$RC_ID) %>%
  merge(data.aud.right, by = c('RC_ID'))

data.rem.aud.ssq.left <- data.merged.ssq %>% filter(RC_ID %in% data.aud.left$RC_ID) %>%
  merge(data.aud.left, by = c('RC_ID')) 


data.rem.aud.ssq.right <- data.rem.aud.ssq.right %>% data.frame('EAR'=c('right'))
data.rem.aud.ssq.left <- data.rem.aud.ssq.left %>% data.frame('EAR'=c('left'))


colnames(data.rem.aud.ssq.left)[which(colnames(data.rem.aud.ssq.left) == 'Age')] <- c('AGE')


data.rem.aud.ssq <- rbind(data.rem.aud.ssq.left,data.rem.aud.ssq.right)  %>% distinct()

data.rem.aud.ssq.plot <- data.rem.aud.ssq  %>% select('AGE','fit.kmeans.cluster','Class','Gender','RC_ID',colnames(data.rem.aud.ssq)[2:18]) %>% na.omit() %>% melt(id = c('RC_ID','fit.kmeans.cluster','Class','AGE','Gender'))

#data.rem.aud.ssq.plot <-  na.omit(data.rem.aud.ssq) %>%  melt(id = c('RC_ID','fit.kmeans.cluster','Class','AGE','Gender'))
data.rem.aud.ssq.plot$value <- data.rem.aud.ssq.plot$value  %>% as.numeric()

png('SSQ_REM_aud.png',width = 2500, 1400)

ggplot(data.rem.aud.ssq.plot, aes(x=variable, y=value, fill = factor(fit.kmeans.cluster))) + 
  geom_boxplot() + xlab('Item #') + ylab('Difference 1st to 2nd visit') + 
  scale_fill_discrete(name = 'Cluster'
                       ) + 
  theme(axis.title.y =  element_text(size = 16),
        axis.title.x = element_text(size = 16),
        legend.title = element_text(size = 16),
        text = element_text(size=12))  + 
  facet_wrap(~Class, ncol = 5)
dev.off()

```

Now do clustering depending on audiogram

```{r}

require('factoextra')

data.aud.cluster <- merge(data.rem1.cluster,data.rem.aud.ssq, by = c('RC_ID'))

freqs <- c('X250','X500','X1000','X2000','X4000','X8000')
audClass <- c('N1','N2','N3','N4','N5','N6','N7','S1','S2','S3')

diffs <- list()
res <- list()
#N1 

for (idx in audClass) {

diffs[[idx]] <- data.aud.cluster %>% filter(Class == idx) %>% select(freqs) 

res[[idx]] <- fviz_nbclust(diffs[[idx]], FUNcluster = kmeans,method = 'wss')

}

```

Now do actual clustering

```{r}

fit.kmeans <- list()

clustNr <- c(3,4,4,3,2,2,2,2,2,3)

p <- list()

cluster <- list()

for (idx in seq(audClass)) {
  
fit.kmeans[[idx]] <-  kmeans(diffs[[idx]],clustNr[idx],nstart = 20)

cluster[[idx]] <- fit.kmeans[[idx]]$centers %>% data.frame(seq(fit.kmeans[[idx]]$size)) %>% data.frame(audClass[idx]) 

}

cluster.plot <- do.call(rbind,cluster)

colnames(cluster.plot) <- c('250','500', '1000', '2000', '4000', '8000','cluster','aud')

png('Cluster_audiogram.png',width = 1200, height = 800)


#now plot the cluster
cluster.plot %>% melt(id=c('cluster','aud')) %>% ggplot(aes(x = variable, y = value, group = factor(cluster), color = factor(cluster))) +
  geom_point(size=4) + 
  geom_line(size=2) +
  xlab('') + ylab('') +
  scale_y_continuous(limits= c(-60,10)) + 
                       xlab('Frequency') + ylab('Difference (dB)') + 
   theme(axis.title.y =  element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 16),
        text = element_text(size = 16),
        legend.position = 'none') + facet_wrap(~ aud)


dev.off()

```

Now connect to SSQ

```{r}

tmp <- list()

for (idx in seq(audClass)){
  
tmp[[idx]] <-  data.aud.cluster %>% filter(Class == audClass[idx]) %>% data.frame('ClusterAud' = fit.kmeans[[idx]]$cluster)
  
}

data.aud.cluster.ssq <- do.call(rbind,tmp) %>% distinct()

# plot SSQ difference per cluster 



```

Now let's connect to the draw users


```{r}

tmp1 <- data.aud.cluster.ssq %>% select(colnames(data.aud.cluster.ssq)[9:25],'RC_ID','ClusterAud','AGE','Gender','Class','EAR')

data.draw.cluster <- t3.data.ssq %>% select('RC_ID','IsT1DrawerUser') %>% merge(tmp1, by = c('RC_ID')) %>% distinct()



```

