---
title: "Drawer Users"
output:
  html_document:
    df_print: paged
---


```{r}
source("setup.R")
```


# 3. Drawer User
  In this section we will attempt to find so-called drawer user, based on the data that we have available. To do this we will first need to define what it means to be a drawer user.

## 3.1 Definition
> A drawer user is ...
>
>*Theories of how to describe a drawer using the data at hand:*
>
- A person who had a baseline meeting/fitting with no follow up meeting
- A person who uses the HA for less than 2 hours a day
- A person who doesnt feel like their HA helps

## 3.2 Motivation

We show that *motivation* has *no significant* influence on hours used.

```{r}

df_motivation_use_group <- merge(df_ha_use, df_motivation, by=c('record_id', 'redcap_event_name', 'own_ha', 'sex','age'))
ggplot(df_motivation_use_group) + geom_point(aes(x=ha_usetime_hours_per_day, y=motivation_line_1_ver2), color="blue") +
               geom_point(aes(x=ha_usetime_hours_per_day, y=motivation_line_2_ver2), color="red") +
               geom_smooth(aes(x=ha_usetime_hours_per_day, y=motivation_line_1_ver2), color="blue") +
               geom_smooth(aes(x=ha_usetime_hours_per_day, y=motivation_line_2_ver2), color="red")

ggplot(df_motivation_use_group) + geom_histogram(aes(x=motivation_line_1_ver2), alpha=0.5, color="blue") +
  geom_histogram(aes(x=motivation_line_2_ver2), alpha=0.5, color="red")

cor(df_motivation_use_group$ha_usetime_hours_per_day, df_motivation_use_group$motivation_line_1_ver2, use = "na.or.complete")
cor(df_motivation_use_group$ha_usetime_hours_per_day, df_motivation_use_group$motivation_line_2_ver2, use = "na.or.complete")

ggplot(df_motivation_use_group)+ 
  geom_histogram(aes(x=motivation_line_1_ver2), alpha=0.5, color="blue")
```

## 3.3 Grouping on usage
We attempt to group observations based on hours of hearing aid use per day. We group them into *low*(<=3 hours), *medium*(<=8 hours) and *high*(<=24 hours) levels.

```{r}
df_use_group <- mutate(df_use_group, use_group = cut(x=ha_usetime_hours_per_day, breaks=c(0,1, 3,8,24), labels = c("Very low","Low", "Medium", "High"), include.lowest = TRUE))

plot(table(Sex=df_use_group$sex, Use=df_use_group$use_group))

ggplot(df_use_group, aes(x=age, y=ha_usetime_hours_per_day)) + geom_point() + geom_smooth()
```

## 3.4 Correlation (SSQ)
We observe no strong correlation between hours of use and any SSQ question. All are negative, which may indicate that; the better hearing the less need for hearing aid.. maybe..

```{r}
library(corrgram)
a <- merge(df_use_group,df_ssq, by=c('record_id', 'redcap_event_name'))
a
corrgram(a[, c(5, 11:23, 25:27)], order = TRUE, 
         main = "Correlogram of SSQ + Use time", 
         lower.panel = panel.conf, upper.panel = panel.ellipse, 
         diag.panel = panel.minmax, text.panel = panel.txt)

a <- merge(df_use_group,ssq_diff, by=c('record_id', 'redcap_event_name'))
a
corrgram(a[, c(5, 11:23)], order = TRUE, 
         main = "Correlogram of SSQ diff + Use time", 
         lower.panel = panel.conf, upper.panel = panel.ellipse, 
         diag.panel = panel.minmax, text.panel = panel.txt)
```

## 3.5 SSQ classification

We attempt to create a *decision tree* based on the SSQ-data, however unsuccesfully since there doesn't seem to be a pattern.

*Base*
```{r}
df_ssq_base_group <- merge(df_ssq_base, df_use_group[,c(1, 11)], by='record_id')
df_ssq_base_group <- filter(df_ssq_base_group, !is.na(use_group))
#df_ssq_base_group <- na.omit(df_ssq_base_group)

set.seed(1234)
ind <- sample(2, nrow(df_ssq_base_group), replace = TRUE, prob=c(0.7, 0.3))
train.data <- df_ssq_base_group[ind == 1, ]

form_ssq_base <- use_group ~ ssq_speech_q1 + ssq_speech_q4 + ssq_speech_q10 + ssq_speech_q11 + ssq_speech_q12 + ssq_space_q6 + ssq_space_q9 + ssq_space_q13 + ssq_sound_q2 + ssq_sound_q7 + ssq_sound_q9 + ssq_sound_q14

tree <- ctree(form_ssq_base, data = train.data)
table(predict(tree), train.data$use_group)

```


-```{r}
df_ssq_ioi_use <- merge(df_ssq_base, df_ioi_usetime, by='record_id')
set.seed(1234)
ind <- sample(2, nrow(df_ssq_ioi_use), replace = TRUE, prob=c(0.7, 0.3))
train.data <- df_ssq_ioi_use[ind == 1, ]

treesf <- as.factor(ioi_ha_1) ~ ssq_sppech_q1 + 
  # ssq_speech_q4 +
  # ssq_speech_q10 +
  # ssq_speech_q11 +
  # ssq_speech_q12 +
  # ssq_speech_q14 +
  # ssq_space_q6 +
  # ssq_space_q9 +
  # ssq_space_q13 +
  ssq_sound_q2 +
  ssq_sound_q7 +
  ssq_sound_q9 +
  ssq_sound_q14 +
  ssq_sound_q15 +
  ssq_sound_q16 +
  ssq_sound_q17 +
  ssq_sound_q18

tree <- ctree(treesf, data = train.data)
plot(tree)
-```

*follow*
-```{r}
df_ssq_follow_group <- merge(df_ssq_follow, df_use_group[,c(1, 11)], by='record_id')
df_ssq_follow_group <- filter(df_ssq_follow_group, !is.na(use_group))
#df_ssq_base_group <- na.omit(df_ssq_base_group)

set.seed(1234)
ind <- sample(2, nrow(df_ssq_follow_group), replace = TRUE, prob=c(0.7, 0.3))
train.data <- df_ssq_follow_group[ind == 1, ]

form_ssq_base <- use_group ~ .-ssq12_dk_complete-record_id-redcap_event_name-use_group

tree <- ctree(form_ssq_base, data = train.data)
table(predict(tree), train.data$use_group)
-```



*Diff*
-```{r}
df_ssq_diff_group <- merge(ssq_diff, df_use_group[,c(1, 11)], by='record_id')
df_ssq_diff_group <- filter(df_ssq_diff_group, !is.na(use_group))
#df_ssq_base_group <- na.omit(df_ssq_base_group)

set.seed(1234)
ind <- sample(2, nrow(df_ssq_diff_group), replace = TRUE, prob=c(0.7, 0.3))
train.data <- df_ssq_diff_group[ind == 1, ]

form_ssq_base <- use_group ~ ssq_sppech_q1 + ssq_speech_q4 + ssq_speech_q10 + ssq_speech_q11 + ssq_speech_q12 + ssq_speech_q14 + ssq_space_q6 + ssq_space_q9 + ssq_space_q13 + ssq_sound_q2 + ssq_sound_q7 + ssq_sound_q9 + ssq_sound_q14 + ssq_sound_q15 + ssq_sound_q16 + ssq_sound_q17 + ssq_sound_q18

tree <- ctree(form_ssq_base, data = train.data)
plot(tree)
table(predict(tree), train.data$use_group)
print(tree)
-```

## 3.6 ioi classification

We CAN find patterns in the IoI classification. We are comparing the use_group and seeing whether or not the patients answers corresponds to their actual use_time from file *"HA_experience_use-time"* 

```{r}
df_ioi_group <- merge(df_ioi, df_use_group[,c(1, 11)], by='record_id')
df_ioi_group <- filter(df_ioi_group, !is.na(use_group))
  #na.omit(df_ssq_base_group)

set.seed(1234)
ind <- sample(2, nrow(df_ioi_group), replace = TRUE, prob=c(0.7, 0.3))
train.data <- df_ioi_group[ind == 1, ]

form_ioi <- use_group ~ ioi_ha_1 + ioi_ha_2 + ioi_ha_3 + ioi_ha_4 + ioi_ha_5 + ioi_ha_6 + ioi_ha_7
form_ioi_no1 <- use_group ~ ioi_ha_2 + ioi_ha_3 + ioi_ha_4 + ioi_ha_5 + ioi_ha_6 + ioi_ha_7

tree_ioi <- ctree(form_ioi, data = train.data)
tree_ioi_no1 <- ctree(form_ioi_no1, data = train.data)
table(predict(tree_ioi), train.data$use_group)
plot(tree_ioi)
plot(tree_ioi_no1)

```

## 3.7 15 D classification

```{r}
df_15d_base_group <- merge(df_15d_base, df_use_group[,c(1, 11)], by='record_id')
df_15d_base_group <- na.omit(df_15d_base_group)

set.seed(1234)
ind <- sample(2, nrow(df_15d_base_group), replace = TRUE, prob=c(0.7, 0.3))
train.data <- df_15d_base_group[ind == 1, ]

form_15d <- use_group ~ fifteen_d_1_move + fifteen_d_2_sight +fifteen_d_3_hear          +fifteen_d_4_breathe+fifteen_d_5_sleep+fifteen_d_6_eat+fifteen_d_7_speak+fifteen_d_8_pee+fifteen_d_9_activity+fifteen_d_10_mental+fifteen_d_11_unplesant+fifteen_d_12_depression+fifteen_d_13_stress+fifteen_d_14_power+fifteen_d_15_sex

tree_15d <- ctree(form_15d, data = train.data)
table(predict(tree_15d), train.data$use_group)
plot(tree_15d)

```




## Fordelingen af audiogram typer over tid de bruger headset om dagen
```{r}
ggplot(merge(df_use_group, df_audiogram, by=c("record_id"))) + geom_bar(aes(x=use_group, fill=Class), position = "fill")

ggplot(merge(df_use_group, df_audiogram, by=c("record_id"))) + geom_bar(aes(x=Class, y = ..count.., fill=use_group), 
                          position = 'fill', stat = 'count', color = 'black')
ggplot(merge(df_use_group, df_audiogram, by=c("record_id"))) + geom_smooth(aes(x=ha_usetime_hours_per_day, y=..count.., color=Class), stat = 'count')
```















